---
title: "Cepaea 2017"
author: 
  - "**Maxime Dahirel** (script author)"
  - Valentin Gaudu
  - Armelle Ansart
date: "13/02/2020"
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

(NB: A "classical" .R script version is also available in the same folder as this file, for readers who'd rather use that. That version is much less commented, but both run the same analyses)


## **Introduction**


(see manuscript for details)


## **Starters and data wrangling**
First, let's load the packages we'll need:

```{r load-packages}
library(here)
library(arm)
library(matrixStats)
library(tidyverse)
library(rstan)
library(tidybayes)
library(bayesplot)
library(brms)
library(patchwork)
rstan_options(auto_write = TRUE)
options(mc.cores = 2)
```

Now we load the raw data files:
```{r load-raw-datafiles}
## Load raw dataset
```




## **Main model**

Now, let's fit our main model (please see description in the methods and supplementary material of the paper).



## **Making figures**

(Fig. 1 is a shell photograph, so we start at Fig.2 for data plots)

### Figure 2


### Figure 3


### Figure 4


## **Supplementary Materials**

We also present a model in Supplementary Materials. 
