---
title: "Cepaea 2017"
author: 
  - "**Maxime Dahirel** (script author)"
  - Valentin Gaudu
  - Armelle Ansart
date: "13/02/2020"
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

(NB: A "classical" .R script version is also available in the same folder as this file, for readers who'd rather use that. That version is much less commented, but both run the same analyses)


## **Introduction**


(see manuscript for details)


## **Starters and data wrangling**
First, let's load the packages we'll need:

```{r load-packages}
library(here)
library(arm)
library(modelr)  #for seq_range
library(matrixStats)
library(tidyverse)
library(rstan)
library(tidybayes)
library(bayesplot)
library(brms)
library(cowplot)
library(QGglmm)
library(patchwork)
rstan_options(auto_write = TRUE)
options(mc.cores = 2)

N_chains <- 4
N_iter <- 20000 ## default number is 2000
N_warmup <- 10000 ## default number is 1000
N_thin <- 1 ## thinning frequency, in case it is needed to keep models at a manageable memory size ( 1: no thinning)
```

Now we load the data files:
```{r load-datafiles}
## Load main dataset

data <- read_csv("./cepaea-2017-dataset.csv")

### to do: add the abundance dataset for the supplementary
```


```{r data-wrangling-censoring}

data<-data %>% 
  mutate(censored_fpt = as.numeric(fpt>=1200),
         censored_bold = as.numeric(boldness >= 1200)
  )
```

```{r data-wrangling-recoding}

data<-data %>% 
  mutate(s_temp = scale(temperature), #centred and scaled
         c_order.temp = order.temp - 2.5, ## center without scaling
         c_order.b = order.b - 1.5, ## same
         c_landscape = -0.5 + as.numeric(landscape == "Marais"), ## same
         ##recoding
         is0b=(band_phenotype=="0b")-mean(band_phenotype=="0b"),
         is3b=(band_phenotype=="3b")-mean(band_phenotype=="3b"),
         is5b=(band_phenotype=="5b")-mean(band_phenotype=="5b")
)

mean_temp <- attr(data$s_temp, "scaled:center")
sd_temp <- attr(data$s_temp, "scaled:scale")
```

```{r data-wrangling-subset}
####add dummy values to be able to use the subset approach without restructuring the table and without
### setting NA flags 
### don't worry, dummy values will be ignored during fitting

data <- data %>% 
  mutate(is.valid.bold = as.numeric(is.na(boldness) == FALSE), ### if it's not NA, it's valid,
         censored_bold2 = c(censored_bold[1:720], rep(1,720)),
         boldness2 = c(boldness[1:720], rep(1200, 720)) 
         ### and we fill the cells that won't be used with dummy values
  )
```


## **Main model**

Now, let's fit our main model (please see description in the methods and supplementary material of the paper).

```{r multivariate-model}
####FORMULAS

bf_explo <- bf(fpt |cens(censored_fpt) ~ ((is3b+is5b) * c_landscape) * s_temp + c_order.temp +
                    (s_temp | p | box) + (s_temp | q | id), family = lognormal)

bf_boldness <- bf(boldness2 | subset(is.valid.bold) + cens(censored_bold2) ~ (is3b+is5b) * c_landscape + c_order.b +
                    (1 | p | box) + (1 | q | id), family = lognormal) ##when sigma not modeled link is identity (half prior?)


### PRIOR
prior_multi <- c(
  set_prior("normal(0, 1)", class = "b",resp=c("boldness2","fpt")), ##### Fixed effects; weakly informative prior for response centred and scaled
  set_prior("normal(log(400),0.5)",class=c("Intercept"),resp=c("boldness2","fpt")),
  set_prior("lkj(2)", class = "cor"),
  set_prior("normal(0,1)",class = "sd",resp=c("boldness2","fpt")),
  set_prior("normal(0,1)",class = "sigma",resp=c("boldness2","fpt"))
)

### fit models ##NB: the sampler may throw initialization errors at the beginning because it evaluates loglik at log(0)
### but then come out fine
###better prior for intercepts?

###FITTING
mod <- brm(mvbf(bf_explo + bf_boldness, rescor = FALSE),
              data = data, chains=N_chains/2, iter = N_iter/100, warmup = N_warmup/100, thin = N_thin, prior = prior_multi, seed = 42,
              control=list(adapt_delta = 0.95,max_treedepth=15) #,save_ranef=FALSE
)

#careful, >1Gb with iter =20000 warmup = 10000 if saving everything
# use save_ranef=FALSE to masssively reduce size (if you don't need individual random effects, simply variances, for inference)
```


## **Making figures**

(Fig. 1 is a set of photographs, so we start at Fig.2 for data plots)

### Figure 2

```{r figure-2}
### PLOTS
##### representing predictions on plots
## at average temperatures (everything else except phenotype is averaged out)

newdata <- data %>% 
  select(band_phenotype,is3b,is5b) %>% 
  distinct() %>% 
  mutate(s_temp = 0, 
         c_order.b = 0, c_order.temp = 0, 
         c_landscape = 0, 
         id = data$id[1], 
         box = data$box[1],
         is.valid.bold=1) %>%
  add_fitted_draws(model = mod, re_formula = NA, scale="linear",resp=c("boldness2")) %>%
  ### scale="response" seems to give the mean on the response scale, but I'd like the median; more representative for a lognormal
  ### so I use scale = "linear" (gives the mean on the log scale, so the median on the response scale)
  ### and I'll back-transform after
  mutate(lboldness = .value) %>% 
  select(-(.value))

newdata2 <- data %>% 
  select(band_phenotype,is3b,is5b) %>% 
  distinct() %>% 
  mutate(s_temp = 0, 
         c_order.b = 0, c_order.temp = 0, 
         c_landscape = 0, 
         id = data$id[1], 
         box = data$box[1],
         is.valid.bold=1) %>%
  add_fitted_draws(model = mod, re_formula = NA, scale = "linear",resp=c("fpt")) %>%
  mutate(lfpt = .value) %>% 
  select(-(.value))

newdata<-inner_join(newdata,newdata2) ; rm(newdata2)

#### to do: need to separate speed and boldness in two tables, add means, backtransform, mise en page
### for boldness
fig_2a <- filter(data,is.na(boldness)==FALSE) %>% 
  ggplot() +
  geom_hline(yintercept=1200,lty=2)+
  geom_line(aes(x = as.numeric(factor(band_phenotype))+0.8*c_order.b, y=boldness, group=id),col = "grey", show.legend = FALSE,alpha=0.4) +
  geom_violin(data=newdata,aes(y = exp(lboldness), x = band_phenotype, fill = band_phenotype),draw_quantiles = 0.5, show.legend = FALSE,alpha=0.8) +
  scale_fill_brewer(palette = "YlOrBr", direction = 1) +
  scale_y_log10(name = "Boldness (latency, sec)", limits = c(3, 1200), breaks = c(5, 10, 20, 50, 100, 200, 500, 1000)) +
  scale_x_discrete(name = "Shell phenotype", labels = c("0 bands", "3 bands", "5 bands")) +
  theme_cowplot()


###arrow_bold
arrow_bold=ggplot()+
  geom_segment(aes(x=c(0,0),xend=c(0,0),y=c(0.05,0.95),yend=c(0.95,0.05)),arrow=arrow())+
  geom_text(aes(x=c(0,0),y=c(0,1),label=c("bolder","shyer")))+theme_void()

###arrow_speed
arrow_speed=ggplot()+
  geom_segment(aes(x=c(0,0),xend=c(0,0),y=c(0.05,0.95),yend=c(0.95,0.05)),arrow=arrow())+
  geom_text(aes(x=c(0,0),y=c(0,1),label=c("faster","slower")))+theme_void()

### for speeds
fig_2b <- ggplot(data = data) + 
  geom_hline(yintercept=1200,lty=2)+
  geom_line(aes(x = as.numeric(factor(band_phenotype))+0.3*s_temp,fpt, group=id),col = "grey", show.legend = FALSE,alpha=0.4) +
  geom_violin(data=newdata, aes(y = exp(lfpt), x = band_phenotype, fill = band_phenotype),draw_quantiles = 0.5, show.legend = FALSE, alpha= 0.8) +
  scale_fill_brewer(palette = "YlOrBr", direction = 1) +
  scale_y_log10(name = "Exploration (first passage time, sec)", limits = c(250, 1200), breaks = c(250, 500, 750, 1000)) +
  scale_x_discrete(name = "Shell phenotype", labels = c("0 bands", "3 bands", "5 bands")) +
  #geom_line(data = tab, aes(x = band_genotype, y = fpt,group=id), position=position_dodge(width = 0.9),col = "grey", show.legend = FALSE,alpha=0.3)+
  theme_cowplot()



#########PLOT 3 
## at temperatures range (everything else is averaged out)
fig_2c <- expand_grid(Temperature = c(15:25), 
              c_order.b = 0, 
              c_order.temp = 0, 
              c_landscape = 0, 
              id = data$id[1], 
              box = data$box[1],
              is3b = 0,
              is5b = 0,
              is.valid.bold=1) %>%
  mutate(s_temp = (Temperature-mean_temp)/sd_temp) %>%
  add_fitted_draws(model = mod, re_formula = NA,scale = "linear",resp="fpt") %>%
  mutate(lfpt = .value) %>% 
ggplot()+
  geom_hline(yintercept=1200,lty=2)+
  geom_line(data = data, aes(x = temperature, y = fpt,group=id), col = "grey", show.legend = FALSE,alpha=0.3) + 
  stat_lineribbon(aes(x = Temperature,y = exp(lfpt)), .width = 0.95,show.legend=FALSE,fill="pink", alpha= 0.8) +
  scale_y_log10(name = "", limits = c(250, 1200),breaks = c(250, 500, 750, 1000)) +
  scale_x_continuous(name = "Test temperature (°C)", limits = c(14, 26)) +
  theme_cowplot()

arrow_bold + fig_2a + plot_spacer() + arrow_speed + fig_2b + fig_2c + plot_layout(nrow = 2, widths= c(1,5,5)) + plot_annotation(tag_levels="A")

```

### Figure 3


### Figure 4


## **Supplementary Materials**

We also present a model in Supplementary Materials. 
