---
title: "Cepaea 2017"
author: 
  - "**Maxime Dahirel** (script author)"
  - Valentin Gaudu
  - Armelle Ansart
date: "13/02/2020"
output: 
  html_document:
    theme: yeti
    toc: TRUE
    toc_float: TRUE
    code_download: TRUE
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

(NB: A "classical" .R script version is also available in the same folder as this file, for readers who'd rather use that. That version is much less commented, but both run the same analyses)


## **Introduction**


(see manuscript for details)


## **Starters and data wrangling**
First, let's load the packages we'll need:

```{r load-packages}
library(here)
library(arm)
library(modelr)  #for seq_range
library(matrixStats)
library(tidyverse)
library(rstan)
library(tidybayes)
library(bayesplot)
library(brms)
library(cowplot)
library(QGglmm)
library(patchwork)
rstan_options(auto_write = TRUE)
options(mc.cores = 2)

N_chains <- 4
N_iter <- 20000 ## default number is 2000
N_warmup <- 10000 ## default number is 1000
N_thin <- 1 ## thinning frequency, in case it is needed to keep models at a manageable memory size ( 1: no thinning)
```

Now we load the data files:
```{r load-datafiles}
## Load main dataset

data0 <- read_csv("./cepaea-2017-dataset.csv")

data_shells0 <- read_csv("./cepaea-2018-shells.csv")

```


```{r data-wrangling-censoring}

data<-data0 %>% 
  mutate(censored_fpt = as.numeric(fpt>=1200),
         censored_bold = as.numeric(boldness >= 1200)
  )
```

```{r data-wrangling-recoding}

data<-data %>% 
  mutate(s_temp = scale(temperature)[,1], #centred and scaled
         c_order.temp = order.temp - 2.5, ## center without scaling
         c_order.b = order.b - 1.5, ## same
         c_landscape = -0.5 + as.numeric(landscape == "Marais"), ## same
         ##recoding
         is0b=(band_phenotype=="0b")-mean(band_phenotype=="0b"),
         is3b=(band_phenotype=="3b")-mean(band_phenotype=="3b"),
         is5b=(band_phenotype=="5b")-mean(band_phenotype=="5b")
)

mean_temp <- attr(scale(data$temperature), "scaled:center")
sd_temp <- attr(scale(data$temperature), "scaled:scale")
```

```{r data-wrangling-subset}
####add dummy values to be able to use the subset approach without restructuring the table and without
### setting NA flags 
### don't worry, dummy values will be ignored during fitting

data <- data %>% 
  mutate(is.valid.bold = as.numeric(is.na(boldness) == FALSE), ### if it's not NA, it's valid,
         censored_bold2 = c(censored_bold[1:720], rep(1,720)),
         boldness2 = c(boldness[1:720], rep(1200, 720)) 
         ### and we fill the cells that won't be used with dummy values
  )
```


## **Main model**

Now, let's fit our main model (please see description in the methods and supplementary material of the paper).

```{r multivariate-model}
####FORMULAS

bf_explo <- bf(fpt |cens(censored_fpt) ~ ((is3b+is5b) * c_landscape) * s_temp + c_order.temp +
                    (s_temp | p | box) + (s_temp | q | id), family = lognormal)

bf_boldness <- bf(boldness2 | subset(is.valid.bold) + cens(censored_bold2) ~ (is3b+is5b) * c_landscape + c_order.b +
                    (1 | p | box) + (1 | q | id), family = lognormal) ##when sigma not modeled link is identity (half prior?)


### PRIOR
prior_multi <- c(
  set_prior("normal(0, 1)", class = "b",resp=c("boldness2","fpt")), ##### Fixed effects; weakly informative prior for response centred and scaled
  set_prior("normal(log(400),0.5)",class=c("Intercept"),resp=c("boldness2","fpt")),
  set_prior("lkj(2)", class = "cor"),
  set_prior("normal(0,1)",class = "sd",resp=c("boldness2","fpt")),
  set_prior("normal(0,1)",class = "sigma",resp=c("boldness2","fpt"))
)

### fit models ##NB: the sampler may throw some initialization errors at the beginning because it falls on an impossible starting value
### but then come out fine

###FITTING
mod <- brm(mvbf(bf_explo + bf_boldness, rescor = FALSE),
              data = data, chains=N_chains/2, iter = N_iter/20, warmup = N_warmup/20, thin = N_thin, prior = prior_multi, seed = 42,
              control=list(adapt_delta = 0.95,max_treedepth=15) #,save_ranef=FALSE
)

#careful, >1Gb with iter =20000 warmup = 10000 if saving everything
# use save_ranef=FALSE to massively reduce size 
#(but that means you can't recreate Fig. 4, which rely on random effects)
```

```{r ppchecks}

summary_mod <- mod %>%
  posterior_samples() %>%
  select(starts_with(c("Intercept", "b_", "sd_", "cor_", "sigma"))) %>%
  pivot_longer(everything()) %>%
  group_by(name) %>%
  mean_hdi()

summary_mod

mcmc_rank_overlay(mod, pars = summary_mod$name)

###harder to find good pp_checks because of censoring, but ribbons seems a-OK
ppc_ribbon(yrep=predict(mod,newdata=data,resp="fpt",summary=FALSE),
           x=rank(predict(mod,resp="fpt")[,1]),
           y=data$fpt,
           prob = 0.5, prob_outer=0.95)

ppc_ribbon(yrep=predict(mod,newdata=data[1:720,],resp="boldness2",summary=FALSE),
           x=rank(predict(mod,resp="boldness2",newdata=data[1:720,])[,1]),
           y=data$boldness2[1:720],
           prob = 0.5, prob_outer=0.95)



```

## **Making figures**

(Fig. 1 is a set of photographs, so we start at Fig.2 for data plots)

### Figure 2

```{r figure2}
##### representing predictions on plots
## at average temperatures (everything else except phenotype is averaged out)

newdata <- data %>% 
  select(band_phenotype,is3b,is5b) %>% 
  distinct() %>% 
  mutate(s_temp = 0, 
         c_order.b = 0, c_order.temp = 0, 
         c_landscape = 0, 
         id = data$id[1], 
         box = data$box[1],
         is.valid.bold=1) %>%
  add_fitted_draws(model = mod, re_formula = NA, scale="linear",resp=c("boldness2")) %>%
  ### scale="response" seems to give the mean on the response scale, but I'd like the median; more representative for a lognormal
  ### so I use scale = "linear" (gives the mean on the log scale, so the median on the response scale)
  ### and I'll back-transform after
  mutate(lboldness = .value) %>% 
  select(-(.value))

newdata2 <- data %>% 
  select(band_phenotype,is3b,is5b) %>% 
  distinct() %>% 
  mutate(s_temp = 0, 
         c_order.b = 0, c_order.temp = 0, 
         c_landscape = 0, 
         id = data$id[1], 
         box = data$box[1],
         is.valid.bold=1) %>%
  add_fitted_draws(model = mod, re_formula = NA, scale = "linear",resp=c("fpt")) %>%
  mutate(lfpt = .value) %>% 
  select(-(.value))

newdata<-inner_join(newdata,newdata2) ; rm(newdata2)

#### to do: need to separate speed and boldness in two tables, add means, backtransform, mise en page
### for boldness
fig_2a <- filter(data,is.na(boldness)==FALSE) %>% 
  ggplot() +
  geom_hline(yintercept=1200,lty=2)+
  geom_line(aes(x = as.numeric(factor(band_phenotype))+0.8*c_order.b, y=boldness, group=id),col = "grey", show.legend = FALSE,alpha=0.4) +
  geom_violin(data=newdata,aes(y = exp(lboldness), x = band_phenotype, fill = band_phenotype),draw_quantiles = 0.5, show.legend = FALSE,alpha=0.8) +
  scale_fill_brewer(palette = "YlOrBr", direction = 1) +
  scale_y_log10(name = "Boldness (latency, sec)", limits = c(3, 1200), breaks = c(5, 10, 20, 50, 100, 200, 500, 1000)) +
  scale_x_discrete(name = "", labels = c("0 bands", "3 bands", "5 bands")) +
  theme_cowplot()+
  ggtitle("A")


###arrow_bold
arrow_bold=ggplot()+
  geom_segment(aes(x=c(0,0),xend=c(0,0),y=c(0.05,0.95),yend=c(0.95,0.05)),arrow=arrow())+
  geom_text(aes(x=c(0,0),y=c(0,1),label=c("bolder","shyer")))+theme_void()

###arrow_speed
arrow_fpt=ggplot()+
  geom_segment(aes(x=c(0,0),xend=c(0,0),y=c(0.05,0.95),yend=c(0.95,0.05)),arrow=arrow())+
  geom_text(aes(x=c(0,0),y=c(0,1),label=c("faster","slower")))+theme_void()

### for speeds
fig_2b <- ggplot(data = data) + 
  geom_hline(yintercept=1200,lty=2)+
  geom_line(aes(x = as.numeric(factor(band_phenotype))+0.3*s_temp,fpt, group=id),col = "grey", show.legend = FALSE,alpha=0.4) +
  geom_violin(data=newdata, aes(y = exp(lfpt), x = band_phenotype, fill = band_phenotype),draw_quantiles = 0.5, show.legend = FALSE, alpha= 0.8) +
  scale_fill_brewer(palette = "YlOrBr", direction = 1) +
  scale_y_log10(name = "Exploration (first passage time, sec)", limits = c(250, 1200), breaks = c(250, 500, 750, 1000)) +
  scale_x_discrete(name = "Shell phenotype", labels = c("0 bands", "3 bands", "5 bands")) +
  #geom_line(data = tab, aes(x = band_genotype, y = fpt,group=id), position=position_dodge(width = 0.9),col = "grey", show.legend = FALSE,alpha=0.3)+
  theme_cowplot()+
  ggtitle("B")


## at temperatures range (everything else is averaged out)
fig_2c <- expand_grid(Temperature = c(15:25), 
              c_order.b = 0, 
              c_order.temp = 0, 
              c_landscape = 0, 
              id = data$id[1], 
              box = data$box[1],
              is3b = 0,
              is5b = 0,
              is.valid.bold=1) %>%
  mutate(s_temp = (Temperature-mean_temp)/sd_temp) %>%
  add_fitted_draws(model = mod, re_formula = NA,scale = "linear",resp="fpt") %>%
  mutate(lfpt = .value) %>% 
ggplot()+
  geom_hline(yintercept=1200,lty=2)+
  geom_line(data = data, aes(x = temperature, y = fpt,group=id), col = "grey", show.legend = FALSE,alpha=0.3) + 
  stat_lineribbon(aes(x = Temperature,y = exp(lfpt)), .width = 0.95,show.legend=FALSE,fill="pink", alpha= 0.8) +
  scale_y_log10(name = "", limits = c(250, 1200),breaks = c(250, 500, 750, 1000)) +
  scale_x_continuous(name = "Test temperature (°C)", limits = c(14, 26),breaks=c(15,20,25)) +
  theme_cowplot()+
  ggtitle("C")

arrow_bold + fig_2a + plot_spacer() + arrow_fpt + fig_2b + fig_2c + plot_layout(nrow = 2, widths= c(1,5,5))

##we annotate (A,B,V) using gg title in each plot rather than plot_annotation (which I'd prefer), because plot_annotation annotates ALL the plots, and here that means even the arrows

```

### Figure 3

```{r latent-varcomps}

### extracting the relevant latent variances is relatively straightforward

### but to get the total VB and VI for the exploration (fpt) variable, 
### we need to account for variation in temperature and in responses to temperature, 
### application of Johnson 2014 MEE (http://onlinelibrary.wiley.com/doi/10.1111/2041-210X.12225/abstract) to get the total random variance of one trait at one obs level (ie including temperature random effect)
### need to create a design matrix for each level:
design_id <- as.matrix(expand_grid(Intercept=rep(1,length(unique(data$id))),s_temp=unique(data$s_temp)))
design_box <- as.matrix(expand.grid(Intercept=rep(1,length(unique(data$box))),s_temp=unique(data$s_temp)))

##now we're good to go

varcomps_fpt_latent <- tibble(
  beta0 = posterior_samples(mod,pars="b_fpt_Intercept")[,1],
  VF= rowVars(posterior_linpred(mod, re_formula = NA,scale = "linear",resp="fpt")),
  VFstate= rowVars(posterior_linpred(mod, 
                                     newdata= data %>% mutate(s_temp=0,c_order.temp=0,c_order.b = 0), 
                                     re_formula = NA,scale = "linear",resp="fpt")),
  ##above: modified newdata because it is the variance of fixed effect assuming individuals only differ in state
  VI = VarCorr(mod,summary=FALSE)$id$cov[,c("fpt_Intercept","fpt_s_temp"),c("fpt_Intercept","fpt_s_temp")] %>% 
    array_tree(.,margin=1) %>% 
    ##apply equation (11) of Johnson 2014 MEE over the posterior
    map(.,~mean(diag(design_id %*% . %*% t(design_id)))) %>% 
    as_vector(),
  VIint= (VarCorr(mod,summary=F)$id$sd[,"fpt_Intercept"])^2,
  VB= VarCorr(mod,summary=FALSE)$box$cov[,c("fpt_Intercept","fpt_s_temp"),c("fpt_Intercept","fpt_s_temp")] %>%
    array_tree(.,margin=1) %>% 
    ##apply equation (11) of Johnson 2014 MEE over the posterior
    map(.,~mean(diag(design_box %*% . %*% t(design_box)))) %>% 
    as_vector(),
  VBint= (VarCorr(mod,summary=F)$box$sd[,"fpt_Intercept"])^2,
  VD=(posterior_samples(mod,pars="sigma_fpt")[,1])^2
  ) %>% 
  mutate(VP = select(.,c(VF,VI,VB,VD)) %>% rowSums() ) %>% 
  mutate(trait="fpt",type="latent")


varcomps_bold_latent <- tibble(
  beta0=posterior_samples(mod,pars="b_boldness2_Intercept")[,1],
  VF=rowVars(posterior_linpred(mod, re_formula = NA,scale = "linear",resp="boldness2")),
  ##checked! works as intended and only provide predictions for the actual boldness values, it discards the dummies added for the multivariate model (see above _ chunks data-wrangling-subset and multivariate-model)
  VFstate = rowVars(posterior_linpred(mod, 
                                      newdata= data %>% mutate(s_temp=0,c_order.temp=0,c_order.b = 0), 
                                      re_formula = NA,scale = "linear", resp="boldness2")),
  ##above: modified newdata because it is the variance of fixed effect assuming individuals only differ in state
  VI = (VarCorr(mod,summary=F)$id$sd[,"boldness2_Intercept"])^2,
  VIint = (VarCorr(mod,summary=F)$id$sd[,"boldness2_Intercept"])^2, ##identical to VI for boldness
  VB = (VarCorr(mod,summary=F)$box$sd[,"boldness2_Intercept"])^2,
  VBint = (VarCorr(mod,summary=F)$box$sd[,"boldness2_Intercept"])^2, ##identical to VB for boldness
  VD = (posterior_samples(mod,pars="sigma_boldness2")[,1])^2
  ) %>% 
  mutate(VP = select(.,c(VF,VI,VB,VD)) %>% rowSums() ) %>% 
  mutate(trait="boldness",type="latent")
```

```{r obs-varcomps}
###Important: VP(obs) will necessarily be < VP(predicted) because censoring artificially limits VP(obs)

QGlognormal<-list(
  inv.link =function(x){exp(x)},
  var.func=function(x){0}, ##no supplementary distributional variance beyond the one in the log residuals (VD)
  d.inv.link=function(x){exp(x)} #derivative of exp is exp
)
##checked; does give _exact_ observed scales and variances for 'simple' lognormal (no fixed effect, no random effect, only residual var)


beta0_obs=NA
VF_obs=NA
VFstate_obs=NA
VI_obs=NA
VIint_obs=NA
VB_obs=NA
VBint_obs=NA
VP_obs=NA
for(i in 1:dim(varcomps_fpt_latent)[1]){
  beta0_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VF[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$mean.obs
  VF_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VF[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VFstate_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VFstate[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VI_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VI[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VIint_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VIint[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VB_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VB[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VBint_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VBint[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VP_obs[i]<-QGicc(mu=varcomps_fpt_latent$beta0[i],var.comp=varcomps_fpt_latent$VP[i],var.p=varcomps_fpt_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
}
VD_obs=VP_obs-(VF_obs+VI_obs+VB_obs)

varcomps_fpt_obs=data.frame(beta0=beta0_obs,
                               VF=VF_obs,VFstate=VFstate_obs,
                               VI=VI_obs,VIint=VIint_obs,
                               VB=VB_obs,VBint=VBint_obs,
                               VD=VD_obs,
                               trait="fpt",
                               type="obs",
                            VP=VP_obs
) 

varcomps_fpt_obs$iter=1:dim(varcomps_fpt_obs)[1]
varcomps_fpt_latent$iter=1:dim(varcomps_fpt_latent)[1]
vcfpt=rbind(varcomps_fpt_obs,varcomps_fpt_latent)


beta0_obs=NA
VF_obs=NA
VFstate_obs=NA
VI_obs=NA
VIint_obs=NA
VB_obs=NA
VBint_obs=NA
VP_obs=NA
for(i in 1:dim(varcomps_bold_latent)[1]){
  beta0_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VF[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$mean.obs
  VF_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VF[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VFstate_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VFstate[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VI_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VI[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VIint_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VIint[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VB_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VB[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VBint_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VBint[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
  VP_obs[i]<-QGicc(mu=varcomps_bold_latent$beta0[i],var.comp=varcomps_bold_latent$VP[i],var.p=varcomps_bold_latent$VP[i],custom.model = QGlognormal,verbose=F)$var.comp.obs
}
VD_obs=VP_obs-(VF_obs+VI_obs+VB_obs)

varcomps_bold_obs=data.frame(beta0=beta0_obs,
                            VF=VF_obs,VFstate=VFstate_obs,
                            VI=VI_obs,VIint=VIint_obs,
                            VB=VB_obs,VBint=VBint_obs,
                            VD=VD_obs,
                            trait="boldness",
                            type="obs",
                            VP=VP_obs
) 

varcomps_bold_obs$iter=1:dim(varcomps_bold_obs)[1]
varcomps_bold_latent$iter=1:dim(varcomps_bold_latent)[1]
vcbold=rbind(varcomps_bold_obs,varcomps_bold_latent)

vc_complete=rbind(vcfpt,vcbold)
```




```{r figure3}

vc_complete %>%
  select(-beta0) %>% 
  gather(key="VC",value="value",VF,VFstate,VI,VIint,VB,VBint,VD,VP)%>% 
  select(-iter)%>% 
  group_by(trait,type,VC) %>% 
  mean_hdi() %>% 
  print(n=Inf)
#sae but for repeatabilities
vc_complete %>% mutate(Ri=VIint/VP,Rstate=(VIint+VFstate)/VP, ratio = VIint/(VIint+VFstate)) %>% 
  gather(key="RPT",value="value",Rstate,Ri,ratio) %>% 
  group_by(trait,type,RPT) %>% select(RPT,value) %>% mean_hdi()

#show relative variance components
vc_complete %>% gather(key="which",value="VC",2:8) %>% filter(type=="obs") %>% 
  mutate(which=fct_relevel(which,c("VF","VFstate","VB","VBint","VI","VIint","VD")) )%>% 
  mutate(trait=fct_recode(trait,Boldness="boldness",Exploration="fpt")) %>% 
  ggplot()+geom_halfeyeh(aes(y=reorder(which,desc(which)),x=100*VC/VP),point_interval = mean_hdi,.width = 0.001,scale="width")+facet_wrap(~trait)+
  scale_x_continuous(name="Variance explained (%, observed scale)", lim=c(0,100))+
  scale_y_discrete(name="Variance component",
                   labels=c(expression(italic(V[D])),
                            expression(italic(V[I(intercept)])),
                            expression(italic(V[I])),
                            expression(italic(V[B(intercept)])),
                            expression(italic(V[B])),
                            expression(italic(V[F(state)])),
                            expression(italic(V[F]))
                            )) +
  theme_cowplot()+
  background_grid()
```


### Figure 4

```{r figure4}

BLUPS=ranef(mod)$id

fig4_main <- tibble(
  bold=BLUPS[,"Estimate","boldness2_Intercept"],
  fpt=BLUPS[,"Estimate","fpt_Intercept"],
  boldCIlow=BLUPS[,"Q2.5","boldness2_Intercept"],
  boldCIhigh=BLUPS[,"Q97.5","boldness2_Intercept"],
  fptCIlow=BLUPS[,"Q2.5","fpt_Intercept"],
  fptCIhigh=BLUPS[,"Q97.5","fpt_Intercept"]) %>% 
  ggplot()+
  geom_segment(aes(x=boldCIlow,xend=boldCIhigh,y=fpt,yend=fpt),col = "grey", show.legend = FALSE,alpha=0.5)+
  geom_segment(aes(x=bold,xend=bold,y=fptCIlow,yend=fptCIhigh),col = "grey", show.legend = FALSE,alpha=0.5)+
  geom_point(aes(x=bold,y=fpt))+
  scale_x_continuous(name = "Boldness BLUP")+
  scale_y_continuous(name = "Exploration BLUP")+
  theme_cowplot()

fig4_inset <- tibble(
  cor = VarCorr(mod,summary=FALSE)$id$cor[,"boldness2_Intercept","fpt_Intercept"]
) %>% 
  ggplot()+
  geom_density(aes(x=cor),fill="white",col="black")+
  scale_x_continuous(name="",limits=c(-1,1),labels=c(-1,"",0,"",1))+
  scale_y_continuous(NULL,breaks=NULL)+
  geom_vline(xintercept = 0,lty=2)+
  theme(axis.line.y=element_blank(), panel.grid = element_blank())

fig4 <- ggdraw(fig4_main)+draw_plot(fig4_inset,x=0.73,y=0.115,height=0.25,width=0.25)

###arrow_bold
arrow_bold=ggplot()+
  geom_segment(aes(x=c(0.05,0.95),xend=c(0.95,0.05),y=c(0),yend=c(0)),arrow=arrow())+
  geom_text(aes(x=c(0,1),y=c(0,0),label=c("bolder","shyer")))+theme_void()

###arrow_speed
arrow_fpt=ggplot()+
  geom_segment(aes(x=c(0,0),xend=c(0,0),y=c(0.05,0.95),yend=c(0.95,0.05)),arrow=arrow())+
  geom_text(aes(x=c(0,0),y=c(0,1),label=c("faster","slower")))+theme_void()

arrow_fpt + fig4 + plot_spacer() + arrow_bold + plot_layout(nrow=2, widths=c(1,10),heights=c(10,1))

```

## **Supplementary Materials**

We also present a model in Supplementary Materials. 

```{r supplementary-model}

data_shells<- data_shells0 %>% 
  mutate(Ntot=rowSums(select(.,N_0bands:N_5bands))) 

mod_S1 <- brm(N_5bands|trials(Ntot)~0+landscape,family=binomial,
      data=data_shells,
      prior=set_prior("normal(0,1.5)",class="b"),
      chains=4,iter=2000,warmup=1000)

data_shells %>% 
  add_fitted_draws(mod_S1) %>% 
  ungroup() %>% select(landscape,.value,Ntot) %>% 
  group_by(landscape) %>%
  mutate(.value=.value/Ntot) %>% 
  median_hdi()

data_shells %>% 
  add_fitted_draws(mod_S1) %>% 
  ungroup() %>% select(landscape,.value,Ntot,.iteration,.draw) %>% 
  mutate(.value=.value/Ntot) %>% 
  compare_levels(variable=.value,by=landscape) %>% 
  median_hdi()


data_shells$y_shells <- with(data_shells, cbind(N_0bands,N_midband,N_2bands,N_3bands,N_5bands))

###works also with a multinomial model if we want to account for all band numbers, and not just 5 vs. the rest
### but defining good priors for "log-odds relative to reference" harder than for "logit (probability)"

mod_S1bis <- brm(y_shells|trials(Ntot)~0+landscape,family=multinomial,
      data=data_shells,
      prior=set_prior("normal(0,1)",class="b",dpar =c("muN2bands","muN3bands","muN5bands","muNmidband")),
      chains=4,iter=2000,warmup=1000)

data_shells %>% 
  add_fitted_draws(mod_S1bis) %>% 
  ungroup() %>% select(landscape,.category,.value,Ntot) %>% 
  group_by(landscape,.category) %>%
  mutate(.value=.value/Ntot) %>% 
  median_hdi(.value)

data_shells %>% 
  add_fitted_draws(mod_S1bis) %>% 
  ungroup() %>% select(landscape,.category,.value,Ntot,.iteration,.draw) %>% 
  group_by(.category) %>% 
  mutate(.value=.value/Ntot) %>% 
  compare_levels(variable=.value,by=landscape) %>% 
  median_hdi()

```

